---
title: "Assignment_5_Adriana"
author: "Adriana Medina"
date: "2024-02-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

# Load packages
```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
```

# Load raw data
```{r}
raw_data_path<-("https://raw.githubusercontent.com/amedina613/Data-607-Week-5-Assignment/main/Data%20607%20week%205%20raw.csv")

raw_data <- read.csv(raw_data_path, header=T)
print(raw_data)
```
# Clean data
My goal is to have a final data set with the following column names: Airline, Destination, On time, and Delayed. 

I will start off by removing that 3rd row. 

```{r}
raw_data <- slice(raw_data, -3)

print(raw_data)
```
# I have some blank values so I'll fill in the blank with the corresponding airline. 

```{r}
names(raw_data)[1:2] <- c("Airline", "Arrival")

raw_data[2, "Airline"] <- raw_data[1, "Airline"]

raw_data[4, "Airline"] <- raw_data[3, "Airline"]

print(raw_data)

```
# We now have a complete data set and no more NA or blank values
I want to transpose the destination information and make a new column called "Destination". I'll use pivot_longer from tidyr for this. 

```{r}
raw_data <- raw_data %>%
  pivot_longer(cols = c(Los.Angeles, Phoenix, San.Diego, San.Francisco, Seattle), names_to = "Destination", values_to = "num_on_time")

print(raw_data)
```
# For the Arrival column I want to pivot that data from long to wide. I will use pivot_wider for this.
```{r}
raw_data<- raw_data %>%
  pivot_wider(names_from = Arrival, values_from = num_on_time)

print(raw_data)
```
# I'm satisfied with the tidying of the data and now we can perform analysis to compare the arrival delays for the two airlines.

```{r}
airline_del <- ggplot(raw_data, aes(x = Airline, y = delayed, fill = Airline)) +
  geom_bar(stat = "identity") +
  labs(title = "Comparison of Delayed Flights for Different Airlines",
       x = "Airline", y = "Number of Delayed Flights") +
  theme_minimal()

print(airline_del)
```


AM West has significantly more flight delays than Alaska Airlines


# Let's see how many flights have arrived on time.
```{r}
airline_ot <- ggplot(raw_data, aes(x = Airline, y = `on time`, fill = Airline)) +
  geom_bar(stat = "identity") +
  labs(title = "Comparison of On Time Flights for Different Airlines",
       x = "Airline", y = "On-time Flights") +
  theme_minimal()

print(airline_ot)
```

AM West has significantly more on time arrivals than Alaska airlines.

# Let's look at a summary of statistics that can better help us determine which airline to avoid.
```{r}
airline_data <- raw_data %>%
  group_by(Airline) %>%
  summarise(
    max = max(delayed),
    min = min(delayed),
    mean = mean(delayed),  
    median = median(delayed),
    standard_deviation = sd(delayed)
  )
print(airline_data)
```

AM west has much greater mean and median of delayed flights than Alaska airlines as well as higher maximum and minimum of delayed flights. 